---
title: "RAVE Pipeline Markdown Template"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
# This code block sets up the engine environment
# Please do not remove me
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
build_pipeline <- raveio::configure_knitr()
# Debug use
# Please set working directory to be in the source file
if(FALSE){
  settings <- raveio::load_yaml('settings.yaml')
  list2env(as.list(settings), envir = globalenv())
#   
  # settings$hclust_method <- "ward.D2"
  # raveio::save_yaml(settings, 'settings.yaml')
}
```



```{rave load_project, language = "R", export = "project"}
library(raveio)
project <- raveio::as_rave_project(project = project_name)
```


```{rave get_input_file_meta_information, language = "R", export = "source_metas"}
# get the exported path of power explorer
search_paths <- project$group_path("power_explorer")

find_source = function(search_paths, fname){
  fpaths = file.path(search_paths, 'exports',fname)
  fexists = file.exists(fpaths)
  if(!any(fexists)){ return(NULL) }
  return(fpaths[which(fexists)[1]])
}

source_metas = lapply(source_files, function(fpath){
  fpath = find_source(search_paths, fpath)
  if( is.null(fpath) ){ return(NULL) }
  dat <- fst::read_fst(fpath, from = 1, to = 1)
  # read.csv( fpath , header = TRUE, nrows = 1)
  list(
    fpath = fpath,
    header = names(dat)
  )
})
source_metas = dipsaus::drop_nulls(source_metas)
```


```{rave load_input_files, language = "R", export = "source_data"}
library(rutabaga)
headers = unique(unlist(lapply(source_metas, '[[', 'header')))
tbls = dipsaus::drop_nulls(lapply(source_metas, function(x){
  
  print('trying to load ' %&% x$fpath)
  
  #progress$inc('Loading...' %&% x$fpath)
  tbl <- fst::read_fst(x$fpath, as.data.table = TRUE)
  # tbl = data.table::fread(file = x$fpath, stringsAsFactors = FALSE, header = TRUE)
  tbl = tbl[tbl$Project %in% project_name, ]
  if(!nrow(tbl)){
    return(NULL)
  }
  mish = headers[!headers %in% names(tbl)]
  for(m in mish){
    tbl[[m]] <- NA
  }
  
  # Load YAML files
  conf = NULL
  yaml_path = paste0(x$fpath, '.yaml')
  if(file.exists(yaml_path)){
    conf = raveio::load_yaml(yaml_path)
  }
  print('returning loaded data ')
  return(list(
    data = tbl,
    conf = conf,
    path = x$fpath,
    subject = tbl$Subject[[1]]
  ))
}))

res = do.call('rbind', lapply(tbls, '[[', 'data'))

if(!is.data.frame(res) || !nrow(res)){
  res = NULL
}else{
  try({
    res$Electrode = as.character(res$Electrode)
    res$Subject = as.character(res$Subject)
    res$Condition = as.character(res$Condition)
  }, silent = TRUE)
  
  subjects = sapply(tbls, '[[', 'subject')
  confs = lapply(tbls, '[[', 'conf')# at here, the lapply is equivalent to do.call(`[[` or "[[",list(tbls,'conf')), or `[[`(tbls,'conf')
  names(confs) = subjects
  
  res = list(
    data = res,
    subjects = subjects,
    confs = confs,
    headers = names(res)
  )
}
source_data <- res

```

```{rave get_ROI_var, language = "R", export = "roi_var"}
# subset with only the selected ROI variable
roi_list <- c("VAR_IS_ROI_Hemisphere", "VAR_IS_ROI_freesurferlabel", 
              "VAR_IS_ROI_Group", "VAR_IS_ROI_Block")
roi_var<- paste0('VAR_IS_ROI_', roi_options$variable)

```

```{rave find_data_column_name, language = "R", export = "var_name"}
var_name = sprintf("%s_%s", baseline_method, analysis_event)
```

```{rave apply_RIO_filters, language = "R", export = "raw_table"}
library(raveclusters)
raw_table <- source_data$data
#var_name = names(raw_table)[3]


roi_list <- c("VAR_IS_ROI_Hemisphere", "VAR_IS_ROI_freesurferlabel", 
              "VAR_IS_ROI_Group", "VAR_IS_ROI_Block")
excluded_roi <- roi_list[!roi_list %in% roi_var]

selected_names <- names(raw_table)
selected_names <- selected_names[!selected_names %in% excluded_roi]

raw_table = raw_table[, !names(raw_table) %in% excluded_roi, with = FALSE]

#select based on the ROI selector
use_regex <- ( roi_options$roi_ignore_gyrus_sulcus || roi_options$roi_ignore_hemisphere )

table_apply_roi <- function(table, roi_column, roi, use_regex){
  var <- table[[roi_column]]
  
  if(use_regex){
    
    pattern <- paste0('(',roi,')', collapse = '|')
    idx <- str_detect(var, pattern)
    
  } else {
    idx <- var %in% roi
  }
  
  return(table[idx,])
}



raw_table <- table_apply_roi(table = raw_table, roi_column = roi_var,
                             roi = filter_by_roi, use_regex = use_regex)


```


```{rave apply_time_filter, languae = "R", export = "collapsed"}
library(rutabaga)
collapsed = lapply(seq_along(input_groups), function( ii ){

  # get the condition groups
  group = input_groups[[ ii ]]
  group_name = group$group_name


  if(is.null(group_name) && group_name == ''){
    group_name = sprintf('Group %d', ii)
  }

  group_condition = group$group_conditions

  # data within plotting time window
  sub_plot = raw_table[raw_table$Condition %in% group_condition &
                         raw_table$Time %within% plot_time_window, ]

  sub_plot$Time = paste0(sub_plot$Time, '_', ii)

  # data within analysis time window
  sub = sub_plot[sub_plot$Time %within% time_window,]

  sub_time = paste0(sub$Time, '_',ii)

  # define the aggregation model
  fml <- Subject + Electrode + VAR_IS_ROI_freesurferlabel ~ Time
  fml[[2]][[3]] <- parse(text = roi_var)[[1]]

  collapsed_mean <- lapply(var_name, function(var){
    reshape2::dcast(
      sub_plot,
      fml,
      fun.aggregate = mean, value.var = var
    )
  }
  )

  merged <- Reduce(function(a, b){
    # b <- collapsed[[1]]
    merge(a, b, all = FALSE,
          by = c("Subject", 'Electrode',roi_var))
  }, collapsed_mean)

  return(list(
    collapsed_mean = merged,
    group_name = group_name,# is this necessary?
    group_index = ii,# and this?
    sub_time = sub_time
  ))
})


```





```{rave data_merging, languae = "R", export = "merged"}
merged = Reduce(function(a, b){
  # b <- collapsed[[1]]
  list(
    collapsed_mean = merge(a$collapsed_mean, b$collapsed_mean, all = FALSE, 
                           by = c("Subject", 'Electrode',roi_var)),
    sub_time = c(a$sub_time, b$sub_time)
  )
}, collapsed, right = FALSE)
```




```{rave data_scaling, languae = "R", export = "baseline"}
library(rutabaga)
baseline = lapply(seq_along(input_groups), function( ii ){
  group = input_groups[[ ii ]]
  
  group_name = group$group_name
  
  if(is.null(group_name) && group_name == ''){
    group_name = sprintf('Group %d', ii)
  }
  
  group_condition = group$group_conditions
  
  baseline_raw = raw_table[raw_table$Condition %in% group_condition &
                             raw_table$Time %within% baseline_time, ]
  
  fml <- Subject + Electrode + VAR_IS_ROI_freesurferlabel ~ Time
  fml[[2]][[3]] <- parse(text = roi_var)[[1]]
  
  baseline <- lapply(var_name, function(var){
    reshape2::dcast(
      baseline_raw,
      fml,
      fun.aggregate = mean, value.var = var
    )
  }
  )
  
  return(baseline)
  
})

baseline_merged = Reduce(function(a, b){
  # b <- collapsed[[1]]
  
  baseline_mean = merge(a, b, all = FALSE, 
                        by = c("Subject", 'Electrode',roi_var))
  baseline_mean
  
}, baseline, right = FALSE)

baseline_mean_indata <- baseline_merged[,!names(baseline_merged) %in% c("Subject", 'Electrode',roi_var)]

baseline_mean <- rowMeans(baseline_mean_indata)
baseline_sd <- apply(baseline_mean_indata,1,sd)

baseline <- list(baseline_mean = baseline_mean, 
                 baseline_sd = baseline_sd)

```



```{rave data_decomposation, languae = "R", export = "collapsed_data"}
collapsed_data = merged$collapsed_mean
if (check_scale) {#with or without 'input', what is the difference? 
  
  collapsed_data[, !names(collapsed_data) %in% c('Subject', 'Electrode', roi_var)] <- 
    t(scale(t(collapsed_data[, !names(collapsed_data) %in% c('Subject', 'Electrode', roi_var)]),
            center = baseline$baseline_mean, baseline$baseline_sd))
}
```


```{rave analysis_data, languae = "R", export = "indata"}

indata = collapsed_data[, !names(collapsed_data) %in% c('Subject', 'Electrode', roi_var)]
if (check_scale) {#with or without 'input', what is the difference? 
  indata = t(scale(t(indata),center = baseline$baseline_mean, 
                   baseline$baseline_sd))
}

```


```{rave measure_distance, language = "R", export = 'dis'}

if(isTRUE(distance_method == '1 - correlation')){
  dis = as.dist(1-cor(t(indata)))
}else if(isTRUE(distance_method == 'DTW')){
  dis = as.dist(dtw::dtwDist(indata))
}else{
  dis = dist(indata, method = distance_method)
}

```


```{rave apply_clustering, language = "R", export = 'clusters'}
n_clust = min(input_nclusters, nrow(indata))

if (input_method == "H-Clust"){
  hcl = stats::hclust(dis, method = hclust_method)
  clusters <- stats::cutree(hcl, k = n_clust)
} else if (input_method == "PAM") {
  km <- cluster::pam(dis, k = n_clust,  cluster.only = TRUE, keep.data = FALSE, 
                     keep.diss = FALSE)
  clusters <- km
}

```


```{rave get_mse, language = "R", export = "mse"}
library(dipsaus)
mse <- lapply(sort(unique(clusters)), function(ci){
  
  apply(collapsed_data[clusters == ci,
                  !names(collapsed_data) %in% c('Subject', 'Electrode', roi_var),
                  drop=FALSE], 
        2, dipsaus::mean_se)
  
})


```


## Build, Visualize, & Run

Please make sure the following code block is at the end of your pipeline file. This block will build the pipeline and generate a `make-rave-builtin-cluster.R` script with your pipeline markdown file. `RAVE` will use the generated pipeline script to execute the pipeline in the dashboard application, or in massive production mode.

```{r build, echo=FALSE, results='hide'}
build_pipeline(make_file = "make-rave-builtin-cluster.R")
```


Once the pipeline script `make-rave-builtin-cluster.R` is built, you can visualize and execute the pipeline without the need of re-knit this document. Notice we use `r` block instead of `rave`. (This is because the code blocks are not part of pipeline targets.)

```{r execute}
Sys.setenv("RAVE_PIPELINE" = normalizePath("."))
raveio::pipeline_run(type = "vanilla")
raveio::pipeline_progress(method = 'details')
```


```{r visualize, echo=FALSE}
Sys.setenv("RAVE_PIPELINE" = normalizePath("."))
raveio::pipeline_visualize()
```








